Formulating RDF Constraints and \\ Validating RDF Data Generically

### 2015-04-22 paper/paper 2/paper 2.tex
evaluation

\section{Introduction}
\tb{feedback: the major problem is semantics. OWL basically work in different semantics than others. And their semantics are formal. Closed world vs. Open World is also a part of this.}
%The purpose of OWL 2 ontologies is to perform reasoning on RDF data and not to validate RDF data conforming to these ontologies.
%OWL 2 is based on the {\em non-unique name assumption} (nUNA) whereas RDF validation requires that different names represent different objects ({\em unique name assumption} (UNA)). 
%Reasoning in OWL 2 is based on the semantics of {\em open-world assumption} (OWA), i.e., a statement cannot be inferred to be false if it cannot be proved to be true  which fits its primary design purpose: to represent knowledge on the World Wide Web. 
%On the other hand, many RDF validation scenarios require the {\em closed-world assumption} (CWA) (i.e., a statement is inferred to be false if it cannot be proved to be true).
%This ambiguity in semantics is one of the main reasons why OWL 2 has not been adopted as a standard constraint language for RDF validation.  
%For this reason, we cannot use OWL 2 ontologies for RDF validation as we use XML Schemas, DTDs, RELAX NG, or Schematron to validate XML documents.

@@

In order to develop a vocabulary to describe constraints of any constraint type generically, it is needed to define the basic terminology for the formulation of constraints and to classify them first. 
A \textbf{\emph{constraint language}} is a language which is used to formulate constraints.
The W3C Data Shapes working group defines \textbf{\emph{RDF constraint}} as a component of a schema what needs to be satisfied\footnote{\url{https://www.w3.org/2014/data-shapes/wiki/Glossary}}.
We identified four dimensions to classify constraints:
\begin{itemize}
  \item \textbf{\emph{Universality:}} \emph{specific constraints} vs. \emph{generic constraints}
	\item \textbf{\emph{Complexity:}} \emph{simple constraints} vs. \emph{complex constraints}
	\item \textbf{\emph{Context:}} \emph{property constraints} vs. \emph{class constraints}
	\item \textbf{\emph{DL Expressivity:}} \emph{constraints expressible in DL} vs. \emph{constraints not expressible in DL}
\end{itemize}

@@

For \emph{property constraints}, \textbf{\emph{left}} and \textbf{\emph{right property lists}} are specified.
The assignment of properties to these lists happens relative to the \emph{constraining element} - which may be an operator (e.g., $\sqsubseteq$ in case of \emph{object property paths}).
\emph{Object Property Paths}\footnote{Corresponds to the requirement \emph{R-55-OBJECT-PROPERTY-PATHS}} (also called \emph{complex role inclusion axioms} in \emph{DL})
express that, if an individual \emph{x} is connected by a sequence of object properties with an individual \emph{y}, 
then \emph{x} is also related to \emph{y} by a particular object property. 
As \emph{Stephen-Hawking} is the \emph{author} of the book \emph{A-Brief-History-Of-Time} (\ms{authorOf(Stephen-Hawking, A-Brief-History-Of-Time)}) whose \emph{genre} is \emph{Popular-Science} ({\ms{genre(A-Brief-History-Of-Time, Popular-Science)}),
the following \emph{object property path} infers that \emph{Stephen-Hawking} is an \emph{author} of the \emph{genre} \emph{Popular-Science} (\ms{authorOfGenre(Stephen-Hawking, Popular-Science)}):
\begin{DL}
authorOf $\circ$ genre $\sqsubseteq$ authorOfGenre 
\end{DL}
Thus, when mapped to the \emph{RDF-CV} (see table \ref{tab:property-constraint-object-property-paths}), the properties \emph{authorOf} and \emph{genre} are on the left side of the \emph{constraining element} \emph{$\sqsubseteq$},
and the property \emph{authorOfGenre} is on the right side:

@@

{\em Literal pattern matching}\footnote{Corresponds to {\em R-44-PATTERN-MATCHING-ON-RDF-LITERALS}} restricts literals to match given patterns.
The following \emph{universal restriction}, e.g., ensures that \emph{books} can only have valid \emph{ISBN} identifiers, 
i.e., strings that match a given regular expression:
\begin{DL}
\ms{Book $\sqsubseteq$ $\forall$ identifier.ISBN} 
\end{DL}
Even though, the restriction of the datatype \emph{ISBN} cannot be expressed in \emph{DL}, \emph{OWL 2 DL} can be used to express the \emph{literal pattern matching} constraint:

@@

\textbf{\emph{Complex constraints}} are composed out of \emph{simple constraints} and/or \emph{complex constraints}.
\emph{Exclusive or} is a logical operation that outputs true whenever both inputs differ (one is true, the other is false).
The \emph{complex constraint} \emph{context-specific exclusive or of property groups}\footnote{Corresponds to R-13-DISJOINT-GROUP-OF-PROPERTIES-CLASS-SPECIFIC} 
restricts individuals of given classes to have properties of only one of multiple mutually exclusive property groups.
\emph{Publications}, e.g., are either identified by an \emph{ISBN} and a \emph{title} (for books) or by an \emph{ISSN} and a \emph{title} (for periodical \emph{publications}), but it should not be possible to assign both identifiers to a given \emph{publication}, which can be expressed in \emph{ShEx} as follows:

@@

There are three forms of \emph{OWL RBox axioms}: \emph{role inclusions}, \emph{equivalence} and \emph{disjointness}. 
\emph{OWL} provides a variety of other axiom types: \emph{role transitivity}, \emph{symmetry}, \emph{asymmetry}, \emph{reflexivity} and \emph{irreflexivity}. 
These axiom types are sometimes considered as basic axiom types in \emph{DL} - using some suggestive notation such as
\ms{Trans(ancestorOf)} to express that the role \emph{ancestorOf} is transitive.
Such axioms, however, are just syntactic sugar - 
all role characteristics can be expressed using the basic features of \emph{DL}.
The \emph{irreflexive object properties} constraint type\footnote{Corresponds to \emph{R-60-IRREFLEXIVE-OBJECT-PROPERTIES}} 
restricts that no individual is connected by a given object property to itself \cite{Kroetzsch2012}.
With the following \emph{irreflexive object property} constraint, e.g., one can state that individuals cannot be authors of themselves:
\begin{DL}
\ms{$\top$ $\sqsubseteq$ $\neg$ $\exists authorOf.Self$}
\end{DL}
When mapped to the \emph{RDF-CV} (see table \ref{tab:irreflexive-object-properties-as-complex-constraints}), the \emph{complex constraint} aggregates three \emph{simple constraints} (one \emph{property} and two \emph{class constraints}):

@@

The \emph{primary key properties}\footnote{Corresponds to \emph{R-226-PRIMARY-KEY-PROPERTIES}} constraint type is often useful to declare a given (datatype) property as the "primary key" of a class, so that a system can enforce uniqueness. 
\emph{Books}, e.g., are uniquely identified by their \emph{ISBN} number, i.e., the property \emph{isbn} is inverse functional \ms{$(\ms{funct } isbn\sp{\overline{\ }})$}, 
which can be represented by the \emph{RDF-CV} in form of a \emph{complex constraint} (see table \ref{tab:primary-key-properties-as-complex-constraints}):

@@

The {\em existential quantification}\footnote{Corresponds to {\em R-86-EXISTENTIAL-QUANTIFICATION-ON-PROPERTIES}}
%\ms{Publication $\equiv$ $\exists$ publisher.Publisher} 
above restricts \emph{publications} to have at least one \emph{publisher} and can be expressed by at least 10 different constraint languages\footnote{OWL 2, SPARQL, ShEx, Bibframe, DSP, ReSh, SPIN, DQTP, Pellet IVC, and Stardog ICV}:

@@

According to the \emph{universal quantification}\footnote{Corresponds to \emph{R-91-UNIVERSAL-QUANTIFICATION-ON-PROPERTIES}} above, \emph{authors} of \emph{publications} must be \emph{persons}.
The SPARQL CONSTRUCT query for the constraint type \emph{universal quantification} is shown below:

@@

We evaluated to which extend the most promising five constraint languages fulfill each of the overall 74 requirements to formulate RDF constraints; each of them corresponding to a constraint type.
We recently published a technical report\footnote{Available at: \url{http://arxiv.org/abs/1501.03933}} (serving as appendix of this paper) in which we explain each requirement (constraint type) in detail and give examples for each (represented by listed constraint languages).
This technical report also contains mappings to DL to logically underpin each requirement and to determine which DL constructs are needed to express each constraint type \cite{BoschNolleAcarEckert2015}.
If a constraint type can be expressed in DL, we added the mapping to DL and to the \emph{generic constraint}.
If a constraint type cannot be expressed in DL, we only added the mapping to the \emph{generic constraint}.
Therefore, we show that each constraint type can be mapped to a \emph{generic constraint}.
Table \ref{tab:evaluation} shows the absolute numbers and the relative percentages for each of the three dimensions to classify constraint types \emph{context}, \emph{complexity}, and \emph{DL expressivity}:

@@

\begin{table}
  %\scriptsize
  %\sffamily
  \vspace{0cm}
	\centering
		\begin{tabular}{l|c|c}
      \textbf{Constraint Classes} & \textbf{\#} & \textbf{\%} \\
      \hline
\emph{Property Constraints} & 48 & 65 (64.86) \\
\emph{Class Constraints} & 17 & 23 (22.96) \\
\emph{Property and Class Constraints} & 9 & 12 (12.16) \\
\hline
\emph{Simple Constraints} & 46 & 62 (62.16) \\
\emph{Simple Constraints} (Syntactic Sugar) & 10 & 14 (13.51) \\
\emph{Complex Constraints} & 18 & 24 (24.32) \\
\hline
\emph{DL Expressible} & 51 & 69 (68.92) \\
\emph{DL Not Expressible} & 23 & 31 (31.08) \\
\hline
Total & 74 & 100 \\
		\end{tabular}
	\caption{Evaluation}
	\label{tab:evaluation}
\end{table}

@@

Constraints types can be classified as \emph{property constraints} and \emph{class constraints} (classification dimension \emph{context}).
Two thirds of the total amount of constraint types are \emph{property constraints} (e.g. \emph{existential} and \emph{universal quantification}), one fifth are \emph{class constraints} (e.g. \emph{allowed values}\footnote{Corresponds to \emph{R-30-ALLOWED-VALUES-FOR-RDF-OBJECTS} and
\emph{R-37-ALLOWED-VALUES-FOR-RDF-LITERALS}} for literals and objects), and approx. 10\% are composed of both \emph{property and class constraints} (e.g. \emph{context-specific exclusive or of property groups}).
Constraint types may be either atomic (\emph{simple constraints}) or created out of \emph{simple} and/or \emph{complex constraints} (\emph{complex constraints}) (classification dimension \emph{complexity}).
Almost two thirds of the constraint types are \emph{simple constraints} (e.g. \emph{object property paths}), a quarter are \emph{complex constraints} (e.g. \emph{ranges of literal values}\footnote{Corresponds to \emph{R-45-RANGES-OF-RDF-LITERAL-VALUES}}), and
nearly 15 percent are \emph{complex constraints} which can be formulated as \emph{simple constraints} when using them in terms of syntactic sugar (e.g. \emph{(un)qualified cardinality restrictions}).
Constraint types can either be \emph{expressible in DL} or \emph{not} (classification dimension \emph{DL expressivity}).
The majority - nearly 70\% - of the overall constraint types are expressible in DL.	
\emph{Literal pattern matching} is an example of a constraint type which cannot be represented in DL.

@@

\tb{how are typical constraints expressed by different constraint languages}

>>>>

\section{Introduction}
\tb{feedback from Erman: the major problem is semantics. OWL basically work in different semantics than others. And their semantics are formal. Closed world vs. Open World is also a part of this.}
%The purpose of OWL 2 ontologies is to perform reasoning on RDF data and not to validate RDF data conforming to these ontologies.
%OWL 2 is based on the {\em non-unique name assumption} (nUNA) whereas RDF validation requires that different names represent different objects ({\em unique name assumption} (UNA)). 
%Reasoning in OWL 2 is based on the semantics of {\em open-world assumption} (OWA), i.e., a statement cannot be inferred to be false if it cannot be proved to be true  which fits its primary design purpose: to represent knowledge on the World Wide Web. 
%On the other hand, many RDF validation scenarios require the {\em closed-world assumption} (CWA) (i.e., a statement is inferred to be false if it cannot be proved to be true).
%This ambiguity in semantics is one of the main reasons why OWL 2 has not been adopted as a standard constraint language for RDF validation.  
%For this reason, we cannot use OWL 2 ontologies for RDF validation as we use XML Schemas, DTDs, RELAX NG, or Schematron to validate XML documents.

@@

In order to develop a vocabulary to describe constraints of any constraint type generically, it is needed to define the basic terminology for the formulation of constraints and to classify them. 
A \textbf{\emph{constraint language}} is a language which is used to formulate constraints.
The W3C Data Shapes working group defines \textbf{\emph{RDF constraint}} as a component of a schema what needs to be satisfied\footnote{\url{https://www.w3.org/2014/data-shapes/wiki/Glossary}}.
We identified four dimensions to classify constraint types:
\begin{itemize}
  \item \textbf{\emph{Universality:}} \emph{specific constraints} vs. \emph{generic constraints}
	\item \textbf{\emph{Complexity:}} \emph{simple constraints} vs. \emph{complex constraints}
	\item \textbf{\emph{Context:}} \emph{property constraints} vs. \emph{class constraints}
	\item \textbf{\emph{DL Expressivity:}} \emph{constraints expressible in DL} vs. \emph{constraints not expressible in DL}
\end{itemize}

@@

For \emph{property constraints}, \textbf{\emph{left}} and \textbf{\emph{right property lists}} are specified.
The assignment of properties to these lists happens relative to the \emph{constraining element} - which may be an operator (e.g., $\sqsubseteq$ in case of \emph{object property paths}).
\emph{Object Property Paths} (\emph{R-55}; also called \emph{complex role inclusion axioms} in \emph{DL})
express that, if an individual \emph{x} is connected by a sequence of object properties with an individual \emph{y}, 
then \emph{x} is also related to \emph{y} by a particular object property. 
As \emph{Stephen-Hawking} is the \emph{author} of the book \emph{A-Brief-History-Of-Time} (\ms{authorOf(Stephen-Hawking, A-Brief-History-Of-Time)}) whose \emph{genre} is \emph{Popular-Science} ({\ms{genre(A-Brief-History-Of-Time, Popular-Science)}),
the following \emph{object property path} infers that \emph{Stephen-Hawking} is an \emph{author} of the \emph{genre} \emph{Popular-Science} (\ms{authorOfGenre(Stephen-Hawking, Popular-Science)}):
\begin{DL}
authorOf $\circ$ genre $\sqsubseteq$ authorOfGenre 
\end{DL}
Thus, when mapped to the \emph{RDF-CV} (see table \ref{tab:property-constraint-object-property-paths}), the properties \emph{authorOf} and \emph{genre} are on the left side of the \emph{constraining element} \emph{$\sqsubseteq$},
and the property \emph{authorOfGenre} is on the right side:

@@

{\em Literal pattern matching} (\emph{R-44}) restricts literals to match given patterns.
The following \emph{universal restriction}, e.g., ensures that \emph{books} can only have valid \emph{ISBN} identifiers, 
i.e., strings that match a given regular expression:
\begin{DL}
\ms{Book $\sqsubseteq$ $\forall$ identifier.ISBN} 
\end{DL}
Even though, the restriction of the datatype \emph{ISBN} cannot be expressed in \emph{DL}, \emph{OWL 2 DL} can be used to express the \emph{literal pattern matching} constraint:

@@

\textbf{\emph{Complex constraints}} are composed out of \emph{simple constraints} and/or \emph{complex constraints}.
\emph{Exclusive or} is a logical operation that outputs true whenever both inputs differ (one is true, the other is false).
The \emph{complex constraint} \emph{context-specific exclusive or of property groups} (\emph{R-13})
restricts individuals of given classes to have properties of only one of multiple mutually exclusive property groups.
\emph{Publications}, e.g., are either identified by an \emph{ISBN} and a \emph{title} (for books) or by an \emph{ISSN} and a \emph{title} (for periodical \emph{publications}), but it should not be possible to assign both identifiers to a given \emph{publication}, which can be expressed in \emph{ShEx} as follows:

@@

There are three forms of \emph{OWL RBox axioms}: \emph{role inclusions}, \emph{equivalence} and \emph{disjointness}. 
\emph{OWL} provides a variety of other axiom types: \emph{role transitivity}, \emph{symmetry}, \emph{asymmetry}, \emph{reflexivity} and \emph{irreflexivity}. 
These axiom types are sometimes considered as basic axiom types in \emph{DL} - using some suggestive notation such as
\ms{Trans(ancestorOf)} to express that the role \emph{ancestorOf} is transitive.
Such axioms, however, are just syntactic sugar - 
all role characteristics can be expressed using the basic features of \emph{DL}.
The \emph{irreflexive object properties} constraint type (\emph{R-60}) 
restricts that no individual is connected by a given object property to itself \cite{Kroetzsch2012}.
With the following \emph{irreflexive object property} constraint, e.g., one can state that individuals cannot be authors of themselves:
\begin{DL}
\ms{$\top$ $\sqsubseteq$ $\neg$ $\exists authorOf.Self$}
\end{DL}
When mapped to the \emph{RDF-CV} (see table \ref{tab:irreflexive-object-properties-as-complex-constraints}), the \emph{complex constraint} aggregates three \emph{simple constraints} (one \emph{property} and two \emph{class constraints}):

@@

The \emph{primary key properties} constraint type (\emph{R-226}) is often useful to declare a given (datatype) property as the "primary key" of a class, so that a system can enforce uniqueness. 
\emph{Books}, e.g., are uniquely identified by their \emph{ISBN} number, i.e., the property \emph{isbn} is inverse functional \ms{$(\ms{funct } isbn\sp{\overline{\ }})$}, 
which can be represented by the \emph{RDF-CV} in form of a \emph{complex constraint} (see table \ref{tab:primary-key-properties-as-complex-constraints}):

@@

The {\em existential quantification} (\emph{R-86})
%\ms{Publication $\equiv$ $\exists$ publisher.Publisher} 
above restricts \emph{publications} to have at least one \emph{publisher} and can be expressed by at least 10 different constraint languages\footnote{OWL 2, SPARQL, ShEx, Bibframe, DSP, ReSh, SPIN, DQTP, Pellet IVC, and Stardog ICV}:

@@

According to the \emph{universal quantification} (\emph{R-91}) above, \emph{authors} of \emph{publications} must be \emph{persons}.
The SPARQL CONSTRUCT query for the constraint type \emph{universal quantification} is shown below:

@@

We evaluated to which extend the five most promising constraint languages on being the standard fulfill each of the overall 74 requirements to formulate RDF constraints; each of them corresponding to a constraint type.
We recently published a technical report\footnote{Available at: \url{http://arxiv.org/abs/1501.03933}} (serving as appendix of this paper) in which we explain each constraint type in detail and give examples for each (represented by different constraint languages).
The technical report also contains mappings to \emph{DL} to logically underpin each constraint type and to determine which \emph{DL} constructs are needed to express each constraint type \cite{BoschNolleAcarEckert2015}.
If a constraint type can be expressed in \emph{DL}, we added the mapping to \emph{DL} and to the \emph{generic constraint} conforming to the \emph{RDF-CV}.
In contrast, if a constraint type cannot be expressed in \emph{DL}, we only added the mapping to the \emph{generic constraint}.
This way, we prove that each constraint type can be mapped to a \emph{generic constraint} and therefore be represented generically by means of the \emph{RDF-CV}.
Table \ref{tab:evaluation} displays the classification of constraint types according to the dimensions \emph{context}, \emph{complexity}, and \emph{DL expressivity}:

@@

\begin{table}
  %\scriptsize
  %\sffamily
  \vspace{0cm}
	\centering
		\begin{tabular}{l|c|c}
      \textbf{Classes of Constraint Types} & \textbf{\#} & \textbf{\%} \\
      \hline
%\textbf{\emph{Context}} & & \\
\emph{Property Constraints} & 48 & 64.86 \\
\emph{Class Constraints} & 17 & 22.96 \\
\emph{Property and Class Constraints} & 9 & 12.16 \\
\hline
%\textbf{\emph{Complexity}} & & \\
\emph{Simple Constraints} & 46 & 62.16 \\
\emph{Simple Constraints} (Syntactic Sugar) & 10 & 13.51 \\
\emph{Complex Constraints} & 18 & 24.32 \\
\hline
%\textbf{\emph{DL Expressivity}} & & \\
\emph{DL Expressible} & 51 & 68.92 \\
\emph{DL Not Expressible} & 23 & 31.08 \\
\hline
Total & 74 & 100 \\
		\end{tabular}
	\caption{Evaluation}
	\label{tab:evaluation}
\end{table}

@@

Constraint types can be classified as \emph{property constraints} and \emph{class constraints} (\emph{context}).
Two thirds of the total amount of constraint types are \emph{property constraints} (e.g., \emph{R-86: existential quantification} and \emph{R-91: universal quantification}), one fifth are \emph{class constraints} (e.g., \emph{R-30/37: allowed values} for objects/literals), and approx. 10\% are composed of both \emph{property and class constraints} (e.g., \emph{R-13: context-specific exclusive or of property groups}).
According to the dimesion \emph{complexity}, constraint types can be either atomic (\emph{simple constraints}) or complex (\emph{complex constraints}), i.e., created out of \emph{simple} and/or \emph{complex constraints}.
Almost two thirds of the constraint types are \emph{simple constraints} (e.g., \emph{R-55: object property paths}), a quarter are \emph{complex constraints} (e.g., \emph{R-45: ranges of literal values}), and
nearly 15 percent are \emph{complex constraints} which can be formulated as \emph{simple constraints} when using them in terms of syntactic sugar (e.g., \emph{R-81/75/82/76/80/74: (un)qualified cardinality restrictions}).
Constraint types can either be \emph{expressible in DL} or \emph{not} (\emph{DL expressivity}).
The majority - nearly 70\% - of the overall constraint types are expressible in \emph{DL}.	
\emph{Literal pattern matching} (\emph{R-44}) is an example of a constraint type which cannot be represented in \emph{DL}, but can also be represented generically by means of the \emph{RDF-CV}.
Altogether, most of the constraint types are \emph{simple constraints on properties} which are \emph{expressible in DL}. 
Thus, the majority of constraint types are directly and relatively easy formulated in form of \emph{simple constraints} which can be mapped to equivalent \emph{DL} constructs.

@@

\tb{1 page / explain how typical constraints are expressed by different constraint languages}
